---
title: "R_HW02"
author: "Odo Luo"
date: "March 15, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework: Performance Assessment

Use both R and Python to answer the following questions:

1. Using the census data set, choose a few meaningful categorical features as predictors and Income as
target.

2. Create train- and test data using a fixed split (use 1/3 for test set).

3. Fit a k-NN-model and a naive Bayes model. Tune k-NN using 10-times CV.

4. Predict the performances on the test set. Create the confusion matrices and compare the two classifiers
in terms of Accuracy, Recall and Precision.

5. Create an ROC-curve for the naive Bayes model. Choose a good threshold, create new predictions
using this threshold on the test set a



# 01 & 02 & 03
```{r}
library(tidyverse)
library(caret)

```

```{r}
census <- read.csv("census.csv",header = TRUE) %>% 
  as_tibble %>% 
  mutate(id=row_number()) %>% 
  mutate(income=as.numeric(ifelse(income==">50K",1,ifelse(income=="<=50K",0,NA)))) %>% 
  mutate_all(function(x) ifelse(x=="?",NA,x)) %>% 
  drop_na()

head(census)
```

## NB

```{r}
library(sjmisc)

nb <- census %>% 
  select(workclass,education,marital.status,occupation,sex,id,income) %>% 
  filter( workclass!="Without.pay" && education!="Preschool")
cols <- c("workclass","education","marital.status","occupation","sex","income")
nb[cols] <- lapply(nb[cols], factor)  

train <- nb %>%  sample_frac(.70) 

test <-  anti_join(nb, train,'id')
train <- train %>% select(-id)
test <- test %>% select(-id)

predictors<- test[1:(length(train)-1)]
target <- test[,ncol(train)] 
predictors
```


```{r}
model = train(income~ .,train,'naive_bayes',trControl=trainControl(method='cv',number=10))
model
```
```{r}
result = predict(model,predictors)
head(result)
```
## KNN

```{r}
knn_data <- census %>% 
  select(age,education.num,hours.per.week,income,id)

knn_data$income <- as.factor(knn_data$income)
knn_train <- knn_data %>%  sample_frac(.70) 

knn_test <-  anti_join(knn_data, knn_train,'id')
knn_train <- knn_train %>% select(-id)
knn_test <- knn_test %>% select(-id)

knn_predictors<- knn_test[1:3]
knn_target <- knn_test[,ncol(knn_train)] 
```

```{r}
knn_model = train(income~ ., knn_train,'knn',trControl=trainControl(method='cv',number=10))
knn_model
```
```{r}
knn_result = predict(knn_model,knn_predictors)
head(knn_result)
```



