---
title: "SVM_R"
author: "Odo Luo"
date: "March 30, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Homework 4

Using R and Python, train SVM regression models using the BostonHousing data (in R:
data("BostonHousing", package = "mlbench"); in Python: from sklearn import datasets;
datasets.load_boston()). Try the linear, one polynomial and the RBF kernel. Tune the parameters.
Compare the models on a test set using the RMSE. Create a scatterplot with the fitted values of the
best model against the true values.

```{r}
library(mlbench)
library(dplyr)
data(BostonHousing)
data<- BostonHousing[,-4]
summary(data)
```

## Train Test Split

```{r}
data <- data %>%  mutate(id=row_number())
train<- sample_frac(data,0.7)
test<- anti_join(data, train,"id")
test <- test %>%  select(-id)
train <- train %>%  select(-id)
```

## Tuning
```{r}
library(e1071)
library(ggplot2)
library(Metrics)

range = 10 ^ (-4:4)
obj = tune.svm(medv ~ crim+ zn +indus  + nox   + rm   +age   + dis +rad +tax+ ptratio    +  b +lstat, data=train,cost=range, gamma = range)
print(obj$best.model)
```
```{r}
obj$best.parameters
```

## Train & Prediction

```{r}
gamma<- 0.1
cost<- 10

linear <- svm(medv ~ crim+ zn +indus  + nox + rm  +age + dis +rad +tax+ ptratio + b +lstat, data=train,kernel="linear"  ,cost=cost,gamma=gamma)
poly <- svm(medv ~ crim+ zn +indus  + nox + rm  +age + dis +rad +tax+ ptratio + b +lstat, data=train,kernel="polynomial",cost=cost,gamma=gamma)
radial <- svm(medv ~ crim+ zn +indus  + nox + rm  +age + dis +rad +tax+ ptratio + b +lstat, data=train,kernel="radial"  ,cost=cost,gamma=gamma)

test_feature <- test[,-13]
test_target <- test[,13]

linear_prediction <- predict(linear,test_feature )
poly_prediction <- predict(poly,test_feature )
radial_prediction <- predict(radial,test_feature )

```

## RSME

```{r}
library(Metrics)
rmse(linear_prediction,test_target)
rmse(poly_prediction,test_target)
rmse(radial_prediction,test_target)
```

## Plots
```{r}
library(ggplot2)

o_data <- test %>%  mutate(type="original") %>% 
  as_tibble

l_data <- test_feature %>% cbind(linear_prediction) %>%  
  mutate(type="linear") %>%  as_tibble %>% 
  mutate(medv=linear_prediction) %>% select(-linear_prediction)

p_data <- test_feature %>% cbind(poly_prediction) %>%  
  mutate(type="poly") %>%  as_tibble %>% 
  mutate(medv=poly_prediction) %>% select(-poly_prediction)

r_data <- test_feature %>% cbind(radial_prediction) %>%  
  mutate(type="radial") %>%  as_tibble %>% 
  mutate(medv=radial_prediction) %>% select(-radial_prediction)

x = o_data %>% bind_rows(l_data) %>% 
  bind_rows(p_data) %>% 
  bind_rows(r_data)

for(n in names(x)){
  p<-ggplot(x,aes_string(x=n,y="medv",color="type")) +
   geom_point()
  plot(p)
}

```



