---
title: "tidyquant_and_datetime_luo"
author: "Odo Luo"
date: "May 13, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 2- Tidyquant and Date/Time 

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(tidyquant)
```

## Description 

Perform the following tasks
Use the tidyquant package:
1. Get the dividend payments of the company.
2. Extract the date of the last dividend payment.
3. Download and show Yahoo prices from 3 days before to 3 days after the last dividend payment.
4. Compare the Close and Adjusted Price and explain their difference

  ** Assigned Company: VW **

### 1. Get the dividend payments of the company.

```{r}
div <- tq_get("VOW.DE", get = "dividends")

div %>% head()
```

### 2. Extract the date of the last dividend payment.

```{r}
ldate<-div %>% filter(date==max(date)) %>% .[["date"]]
ldate
```


### 3. Download and show Yahoo prices from 3 days before to 3 days after the last dividend payment.

```{r}
stocks <- tq_get("VOW.DE", get = "stock.price", 
                from = as.Date(ldate) - 3, 
                to = as.Date(ldate) + 3)
stocks %>% head()
```


### 4. Compare the Close and Adjusted Price and explain their difference

```{r}
stocks %>% 
  select(symbol,date,close,adjusted) %>% 
  mutate(difference=close-adjusted)
```
Der Adjusted Preis ist der Aktienwert nach Abzug der Dividende.
