---
title: "xts_excercise_luo"
author: "Odo Luo"
date: "May 28, 2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(quantmod)
library(ggplot2)
library(tidyverse)
library(dplyr)
```

# XTS Exercise 

Use the quantmod package:
1. Get the dividend payments of the company.
2. Extract the date of the last dividend payment.
3. Download and show Yahoo prices from 3 days before to 3 days after the last dividend payment.
4. Compare the Close and Adjusted Price and explain their difference


## 1. Get the dividend payments of the company.
```{r}
div<-getDividends("VOW.DE", return.class="ts")
div %>% head()
```

## 2. Extract the date of the last dividend payment.
```{r}
end(div)
```

3. Download and show Yahoo prices from 3 days before to 3 days after the last dividend payment.

```{r}
getSymbols("VOW.DE",from=as.Date(end(div))-3, to=as.Date(end(div))+3)
VOW.DE[,"VOW.DE.Close"] %>%  plot()
```

## 4. Compare the Close and Adjusted Price and explain their difference

```{r}
VOW.DE %>% as.data.frame() %>%  mutate(difference= VOW.DE.Close - VOW.DE.Adjusted)
```


The difference reprecents the dividend VW payed per share. The adjusted price is the close price substracted by that.
