---
title: "The NaiveBayes-Classifier"
author: "David Meyer"
date: "11.10.2018"
output:
  pdf_document: 
    latex_engine: xelatex
    number_sections: yes
    toc: yes
  html_notebook: default
---

```{r}
library(tidyverse)
library(e1071)
```
---

# Data Management

Census Data on yearly income of US citizens:

```{r}
census <- read.csv("census.csv", stringsAsFactors = TRUE)
glimpse(census)
```

# Probability-based classification

Idea for classifier based on probabilities:

Compute P(y|x) for all classes of y. Choose the class with highest probability.

Q: Income class for women?

```{r}
## A-priori P
census %>% count(income) %>% mutate(p = n/sum(n))
```
```{r}
## P, given sex ("A-posteriori-Prob. of Income")
census %>% group_by(sex) %>% count(income) %>% mutate(p = n/sum(n))
```

**Note: the probability for income <= 50K does increase when the algorithm knows that gender is female ("a-posteriori probability"), compared to the *a-priori*-probabilities.**

For both sexes, the algorithm would predict that the income is less than 50K (with a higher probability for women).

Q: Income class for black women?

```{r}
## P, given sex and race
census %>% group_by(sex, race) %>% count(income) %>% mutate(p = n/sum(n))
```

Same predicted class, but black women have an even higher probability for Income < 50K than white women.

*Issues:*

- Computation time increases rapidly with amount of predictors used
- It is unlikely that all combinations have been observed
- How to deal with numeric predictors?

# Naïve Bayes Classifier

Use Bayes Formula:

$$P(\mathrm{income}|\mathrm{sex}) = \frac{P(\mathrm{sex}|\mathrm{income})P(\mathrm{income})}{P(\mathrm{sex})}$$

This is computed for both income classes, given sex = Female, and then we choose the class with highest probability.

Note that the denominator needs not to be computed, since it is the same for income >= 50K and income < 50K. If needed, it computes as the sum of both nominators:

$$\begin{array}{rcl}
P(\mathrm{sex} = \mathrm{Female}) &=& P(\mathrm{sex=Female}|\mathrm{income}=''\ge50K'')P(\mathrm{income}=''\ge50K'')\\ 
&+& P(\mathrm{sex=Female}|\mathrm{income}=''<50K'')P(\mathrm{income}=''<50K'')
\end{array}$$

```{r}
library(e1071)
m = naiveBayes(income ~ sex, data = census)
m
```

```{r}
# P(Female | Y > 50K) * P(Y > 50K)
0.15 * 0.24

# P(Female | Y <= 50K) * P(Y <= 50K)
0.39 * 0.76

# P(Female)
0.15 * 0.24 + 0.39 * 0.76

# P(Y > 50K | Female)
0.036 / 0.3324

# P(Y <= 50K | Female)
0.2964 / 0.3324

predict(m, data.frame(sex = "Female"))
predict(m, data.frame(sex = "Female"), type = "raw")
```

This extends to more than one predictor *assuming that all predictors are independent*:

$$P(\mathrm{income}|\mathrm{sex, race}) = \frac{P(\mathrm{sex}|\mathrm{income})P(\mathrm{race}|\mathrm{income})P(\mathrm{income})}{P(\mathrm{sex, race})}$$


```{r}
m2 =naiveBayes(income ~ sex + race, data = census)
m2
```

```{r}
predict(m2, data.frame(sex = "Female", race = "Black"), type = "raw")
predict(m2, data.frame(sex = "Female", race = "Black"))
```

```{r}
0.11 * 0.388 * 0.76 ## Black/Female for <=50K
0.049 * 0.15 * 0.24 ## Black/Female for >50K
0.11 * 0.388 * 0.76 + 0.049 * 0.15 * 0.24 ## denominator
0.032 / 0.0342 ## P(income <= 50K|Black, Female)
0.001764 / 0.034 ## P(income > 50K|Black, Female)
## --> predict income <= 50K
```

Note that the computed probabilities are not exact---the reason is, that gender is not homogeneously distributed over race, and thus the independence assumption is violated:

```{r}
census %>% group_by(race) %>% count(sex) %>% mutate(p = n/sum(n))
```

# metric predictors

In case of metric predictors, NaiveBayes assumes a normal distribution with mean and standard deviation estimated from the sample data. In the formula, the conditional probabilities (e.g., $P(sex|gender$)) are then replaced with the density function.

```{r}
m = naiveBayes(income ~ sex + age, data = census)
m
```

The first column of the ``age'' table stores the means, the second one the standard deviations for age (given Y, i.e. income).

```{r}
predict(m, data.frame(sex = "Female", age = 40), type = "raw")
```

```{r}
a = 0.388 * dnorm(40, 36.78, 14) * 0.76 
b = 0.15 * dnorm(40, 44.25, 10.5) * 0.24
a / (a + b)
b / (a + b)
```

# Exercise

Using the `PimaIndiansDiabetes2` data from package `mlbench` (remove instances with missing data), predict the first 5 instances using a NaiveBayes model.

